---
title: "NBA_Player_value_predictor"
author: "Alexander Nicholls"
date: "31/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(tidyverse)
library(hablar)
library(corrplot)
library(measurements)
library(rpart)
library(randomForest)
library(olsrr)
setwd("C:/Users/alexa/Documents/GitHub/anic0077.github.io/NBA_stats")
```
player_salary
```{r}
salary_2021 <- read.csv("Salary_2021.csv")

```

2021_regular_data
```{r}

regular_traditional2021<- read.csv("2021_regular_traditional.csv")
regular_traditional2021$TEAM <- as.factor(regular_traditional2021$TEAM)
names(regular_traditional2021)[1] <- "name"
```
5 year regular data
```{r}
combined_regular <- read.csv("regular_combined_5_traditional.csv")
drops <- c("TEAM")
combined_regular <- combined_regular[ , !(names(combined_regular) %in% drops)]
agg_regular = aggregate(combined_regular,
                by = list(combined_regular$PLAYER),
                FUN = mean)
drops <- c("PLAYER")
agg_regular <- agg_regular[ , !(names(agg_regular) %in% drops)]
names(agg_regular)[1] <- "name"
```


2021_playoff_data
```{r}
playoff_traditional2021<- read.csv("2021_playoff_traditional.csv")
playoff_traditional2021$TEAM <- as.factor(playoff_traditional2021$TEAM)
#editing the column names
colnames(playoff_traditional2021) <- paste("Playoff", colnames(playoff_traditional2021), sep = "_")
names(playoff_traditional2021)[1] <- "name"
#names(playoff_traditional2021)[2] <- "TEAM"
#names(playoff_traditional2021)[3] <- "AGE"
```
5 year playoff data
```{r}
combined_playoffs <- read.csv("playoff_combined_5_traditional.csv")
drops <- c("TEAM")
combined_playoffs <- combined_playoffs[ , !(names(combined_playoffs) %in% drops)]

agg_playoffs = aggregate(combined_playoffs,
                by = list(combined_playoffs$PLAYER),
                FUN = mean)
drops <- c("PLAYER")
agg_playoffs <- agg_playoffs[ , !(names(agg_playoffs) %in% drops)]
colnames(agg_playoffs) <- paste("Playoff", colnames(agg_playoffs), sep = "_")
names(agg_playoffs)[1] <- "name"
```

Merging datasets
```{r}
#merging regular and salary
Nba_2021 <- merge(regular_traditional2021,salary_2021,by="name")
#merging regular and playoff
NBA_2021 <- merge(Nba_2021,playoff_traditional2021, by='name')
```
merging 5 year datasets
```{r}
Nba_2021 <- merge(agg_regular,salary_2021,by="name")
NBA_2021 <- merge(Nba_2021,agg_playoffs, by='name')
```

predicting salary
```{r}
model <- lm(salary_2021 ~ . - name - Playoff_AGE - AGE - FP - Playoff_FP, data = NBA_2021)
#ols_step_forward_p(lin)
step <- ols_step_both_p(model)
step
```
constructing the optimal model
```{r}
NBA_salary <- lm(salary_2021 ~ PTS + X... + Playoff_STL + Playoff_DREB     + PF + AST + TOV + FT. + Playoff_L + Playoff_X... + PF + Playoff_OREB, data = NBA_2021)
summary(NBA_salary)

#NBA_salary <- lm(salary_2021 ~ FP + PF + X... + Playoff_FGA + Playoff_STL + Playoff_X... + TOV + Playoff_L + BLK + FT., data = NBA_2021)
#summary(NBA_salary)
```
creates table of players value
```{r}
value <- data.frame(name=character(), value=numeric(), is_paid = numeric(), is_worth=numeric())
index = 1
Nba_2021_nd <-  NBA_2021[!duplicated(NBA_2021$name), ]
for (i in Nba_2021_nd$name){
  #print(i)
  value[index, "name"] <- i
  test_data <- Nba_2021_nd[Nba_2021_nd$name == i,]
  prediction <- predict.lm(NBA_salary, test_data, interval = "confidence", type = "response")
  #print(prediction[1] - test_data$salary_2021)
  value[index, "value"] <-(prediction[1] - test_data$salary_2021)
  value[index, "is_paid"] <- test_data$salary_2021
  value[index, "is_worth"] <- prediction[1]
  index <- index + 1
}
```
Plotting above
```{r}
plot(value$is_paid, value$is_worth)
abline(0,1)
```
trying out a cubist model
```{r}
library(Cubist)

# Fit the model, note different syntax in calling the function
cubist.model <- cubist(x= NBA_2021[,c(3:24, 26:28, 31:53, 54:56)], y=NBA_2021[,29])


summary(cubist.model) # Print out the model

cubist.predict<-predict(cubist.model, newdata = NBA_2021[,c(3:24, 26:28, 31:53, 54:56)])
plot(NBA_2021$salary_2021,cubist.predict,xlab="observed",ylab="predicted")
abline(0,1)
```
create table of value for cubic model
```{r}
value_cubic <- data.frame(name=character(), value=numeric(), is_paid = numeric(), is_worth=numeric())
index = 1
Nba_2021_nd <-  NBA_2021[!duplicated(NBA_2021$name), ]
for (i in Nba_2021_nd$name){
  #print(i)
  value_cubic[index, "name"] <- i
  test_data <- Nba_2021_nd[Nba_2021_nd$name == i,]
  prediction <- predict(cubist.model, test_data, interval = "confidence", type = "response")
  #print(prediction[1] - test_data$salary_2021)
  value_cubic[index, "value"] <-(prediction[1] - test_data$salary_2021)
  value_cubic[index, "is_paid"] <- test_data$salary_2021
  value_cubic[index, "is_worth"] <- prediction[1]
  index <- index + 1
}
```

Random forest model
```{r}
# Random Forest
library(randomForest)
## Part i - Model fitting

#ntree = number of bootstrap samples
#(salary_2021 ~ . - name - Playoff_AGE - AGE - FP - Playoff_FP, data = NBA_2021, ntree = 500, importance = TRUE,do.trace = 10, proximity=TRUE)

rf.model <- randomForest(salary_2021 ~ . - name - Playoff_AGE - AGE - FP - Playoff_FP, data = NBA_2021, ntree = 500, importance = TRUE,do.trace = 10, proximity=TRUE)

plot(rf.model)

round(importance(rf.model), 2)

varImpPlot(rf.model)


par(mfrow=c(2,2))
#partialPlot(rf.model,soil,pH)
#partialPlot(rf.model,soil,Elevation)
#partialPlot(rf.model,soil,Gamma.K)
#partialPlot(rf.model,soil,Valley_Bottom)


rf.inb.predict<-predict(rf.model,NBA_2021)
par(mfrow=c(1,1))
plot(NBA_2021$salary_2021,rf.inb.predict,xlab="observed",ylab="in-bag prediction")
abline(0,1)


rf.oob.predict<-rf.model$predicted
plot(NBA_2021$salary_2021,rf.oob.predict,xlab="observed",ylab="out-of-bag prediction")
abline(0,1)
```
table for rf values
```{r}
value_rf <- data.frame(name=character(), value=numeric(), is_paid = numeric(), is_worth=numeric())
index = 1
Nba_2021_nd <-  NBA_2021[!duplicated(NBA_2021$name), ]
for (i in Nba_2021_nd$name){
  #print(i)
  value_rf[index, "name"] <- i
  test_data <- Nba_2021_nd[Nba_2021_nd$name == i,]
  prediction <- predict(rf.model, test_data, interval = "confidence", type = "response")
  #print(prediction[1] - test_data$salary_2021)
  value_rf[index, "value"] <-(prediction[1] - test_data$salary_2021)
  value_rf[index, "is_paid"] <- test_data$salary_2021
  value_rf[index, "is_worth"] <- prediction[1]
  index <- index + 1
}
```

assessing the models
```{r}
accuracy <-function (actual,predict){
  # Calculate RMSE and ME
  rmse=sqrt(mean((actual-predict)^2))
  me=mean(actual-predict)
  # Calculate R2
  sse=sum((actual-predict)^2)
  sst=sum((actual-mean(actual))^2)
  R2=1-sse/sst
  # Calculate Concordance Correlation (rho, p)
  k <- length(actual)
  xb <- mean(actual)
  yb <- mean(predict)
  sx2 <- 1/k * sst
  sy2 <- 1/k * sum((predict-yb)^2)
  sxy <- 1/k * sum((actual-xb)*(predict-yb))
  p <- 2 * sxy/(sx2 + sy2 + (yb - xb)^2)
  # print our values    
  cat("ME=", me, " RMSE=", rmse, " R2=", R2, " rho.c= ", p)
  return(c(me,rmse,R2,p))}
```

```{r}
r_forest<-accuracy(NBA_2021$salary_2021,rf.oob.predict)
r_cubic<-accuracy(NBA_2021$salary_2021,cubist.predict)
r_linear <-accuracy(NBA_2021$salary_2021, )
```
combining the results of the three classifiers
```{r, warning=FALSE}
value_combined <- rbind(value,value_cubic,value_rf)
value_agg = aggregate(value_combined,
                by = list(value_combined$name),
                FUN = mean)
drops <- c("name")
value_agg <- value_agg[ , !(names(value_agg) %in% drops)]
names(value_agg)[1] <- "name"
```

plotting above
```{r}
plot(value_agg$is_paid, value_agg$is_worth)
abline(0, 1)
```

